---
title: "Welcome Test  1388371"
author: "Mele Umberto Junior"
date: "Thursday, February 23, 2017"
output: html_document
---
<h4 style='color:blue' > 0)  Do you know how to use an RMarkdown file to answer these questions with text and R code?   

```{r, echo=FALSE}
cat('Yes, I Know!')
```

1) Fix the seed to start the simulation of a single random variable from a (standard) uniform distribution on $[0,1]$. Use your matricola number as seed. Call the simulated value $lambda$

```{r, echo=TRUE}
set.seed(1388371)
lambda = runif(1)
lambda
```

2) Simulate 1000 random variables using the following code


```{r}
X = runif(1000)
Y = - lambda * log(X)
hist(Y, freq=FALSE)
```


3) Can you show that the theoretical distribution of Y is exponential with rate equal to $\frac{1}{lambda}$? [explain or . . . do it!]
4) Can you plot a graphical representation of the theoretical distribution ?

```{r echo=FALSE}
cat('Yes, of course!!!')

```

```{r echo=TRUE}
hist(Y, freq=FALSE)
curve(dexp(x, rate = 1/lambda), add=TRUE, col='blue')
```

5) Can you tell what is the relation between the theoretical curve of the density of Y and the graph of the histogram?

```{r}

```




