---
title: 'Homework #0 (aka Welcome Test)'
author: |
  | Stat4DS2+DS (key: `tardella2017`)
  | <https://elearning2.uniroma1.it/course/view.php?id=4951>
date: '**deadline 27/02/2017 (8:30)**'
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Your Last+First Name ______________ Your Matricola _________

0. Do you know how to use an RMarkdown file to answer these questions with text and R code?
```{r}
#
#
#
```


1. Fix the *seed* to start the simulation of a single random variable from a (standard) uniform distribution on [0,1]. Use your *matricola* number as seed. 
Call the simulated value `lambda`
```{r,echo=FALSE,results='hide'}
set.seed(123)
lambda  = runif(1)
```
```{r}
#
#
#
```

2. Simulate 1000 random variables using the following code
```{r}
X = runif(1000)
Y = - lambda * log(X)
```
and then plot the histogram of the simulated values using the option `freq=FALSE`. 

3. Can you **prove** that the theoretical distribution of $Y$ is exponential with rate equal to `1/lambda`? [explain or ... do it with all details!]

```{r}
#
#
#
#
#
#
#
#
#
#
#
#
```

4. Can you plot a graphical representation of the theoretical distribution? 

```{r}
#
#
#
#
#
#
#
#
#
```

5. Can you tell what is the relation between the theoretical curve of the density of $Y$ and the graph of the histogram?

```{r}
#
#
#
#
#
#
#
#
#
#
```


6. Suppose we have a random vector $(Z,Y)$ Do you know what a joint density $f_{Z,Y}(z,y)$ is?  What are the properties it must satisfy? 

```{r}
#
#
#
#
#
#
#
#
#
```

7. Can you derive in formula the distribution of $Z$? How it is called? 

```{r}
#
#
#
#
#
#
#
#
#
#
```

8. Conditionally on each simulated $Y_i=y_i$ (i=1,2,...,1000), simulate $Z_i = y_i*W_i$ where $W_i$ has distribution standard Normal and is independent of Y_i. What is the conditional distribution of $Z|Y=y$?

```{r}
#
#
#
#
#
#
#
#
#
#
#
#
#
#
```

9. Can you provide a simulation from the marginal distribution of $Z$? 

```{r}
#
#
#
#
#
#
#
#
#
#
```

10. Can you provide an approximation of the marginal distribution of $Z$? Please justify your answer. 

```{r}
#
#
#
#
#
#
#
#
#
#
```

11. How could you derive/write the joint distribution of $(Z,Y)$? 

```{r}
#
#
#
#
#
#
#
#
#
#
```

* * *
<div class="footer"> &copy; 2016-2017 - Stat4DS2+CS - Luca Tardella </div>

```{r, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
cat(paste0("This homework will be graded and it will be part of your final evaluation \n\n "))
cat(paste("Last update by LT:",date()))
```
